---
title: "R Notebook - Read all Excel sheets and store them into separate tibbles (with SQL)"
date: "March 2023"
author: "Daan Jansen"
output:
  html_document:
    df_print: paged
    html_notebook: default
    pdf_document: default
    theme:
      bootswatch: journal
---

\

***
We start with an extensive setup chunk in which we:

1. *Load* all the *packages* we use;
2. *Read* the *data* from a SQL Server database and load tables into separate tibbles
3. *Clean data* and preparation
\

***
\

### 1. Load libraries
```{r Loading libraries, include=FALSE}

# Packages have proprogrammed functions that solve specific problems and expands the R vocabulary
library("tidyverse") # includes packages like dplyr and ggplot2
library("readxl") # library to read Excel files
library("DBI") # Ephemeral in-memory RSQLite database
library("lubridate") # working with dates and times

```
`End load libraries`

<BR><BR>

### 2. Read all Excel sheets and load into separate tibbles
```{r Loading data from Excel, include=FALSE}

# Change these parameters accordingly
path <- "data/KlantArtikel.xlsx" # change file name accordingly
endsWith <- "datum"

# Create an in-memory SQLite database
con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
# Helper function to convert date columns for SQLite usage
convertDateColumns <- function(x, endsWith) {
  x %>% 
    mutate_at(vars(ends_with(endsWith)), as.character)
}

# Read all Excel sheets
sheets <-
  excel_sheets(path) %>%
  set_names() %>%
  map(read_excel, path = path)

# Loading sheets into tables in SQLite database
  lapply(seq_along(sheets), function(x) {copy_to(con, convertDateColumns(sheets[[x]], endsWith), names(sheets)[[x]])})
# Expanding sheets list to separate tibbles
  lapply(seq_along(sheets), function(x) {assign(names(sheets)[[x]], sheets[[x]], envir = sys.frame())})
  
# Clean up
rm(sheets, convertDateColumns, path, endsWith)

```
`End read and load data`

<BR><BR>

### 3. Data cleaning and preparation
```{r data cleaning and preparation}

# OPTIONAL: Remove empty observations (rows.)
# cleandf <- drop_na(exceldf, <NAME>) # Delete observations with the value NA after a specific variable
# cleandf <- na.omit(exceldf) # Remove all observations where a variable has the value NA

#knitr::opts_chunk$set(warning = FALSE, message = FALSE) # suppress all warning messages 
```
`End data cleaning and preparation`

\

***
**--> Example chunks <--**

***
\

### Example SQL code chunk using dbGetQuery
```{r Example histogram from dbGetQuery}

# Returns the result of a SQL query as a data frame
query1 <- dbGetQuery(con,
                     "SELECT catcode, count(catcode) as aantal
                     FROM artikel
                     GROUP BY catcode")

query1 %>% 
  ggplot(aes(x = catcode, y = aantal)) + 
  geom_col()

```
`End example SQL code chunk using dbGetQuery`

<BR><BR>

### Example SQL code chunk capture in R dataframe and plot result
```{sql connection=con, output.var = "query1", label='Store result in outputvariable'}

SELECT catcode, count(catcode) as aantal
FROM artikel
GROUP BY catcode

```

```{r Plot result from previous SQL chunk}

query1 %>% 
  ggplot(aes(x = catcode, y = aantal)) + 
  geom_col()

```
`End example SQL code chunk capture in R dataframe and plot result`

<BR><BR>

### Example code chunk from dataframe
```{r Example}

  artikel %>% 
  select(catcode) %>% 
  ggplot(aes(catcode)) +
  geom_histogram(stat = "count")

```
`End example code chunk from dataframe`

\

***
**--> From here you make your own chunks <--**

***
\
