---
title: "R Notebook - Read one Excel sheet"
date: "March 2023"
author: "Daan Jansen"
output:
  html_document:
    df_print: paged
    html_notebook: default
    pdf_document: default
    theme:
      bootswatch: journal
---

\

***
We start with an extensive setup chunk in which we:

1. *Load* all the *packages* we use;
2. *Read* the *data* from a SQL Server database and load tables into separate tibbles
3. *Clean data* and preparation
\

***
\

### 1. Load libraries
```{r Loading libraries}

# Packages have proprogrammed functions that solve specific problems and expands the R vocabulary
library("tidyverse") # includes packages like dplyr and ggplot2
library("readxl") # library to read Excel files

```
`End load libraries`

<BR><BR>

### 2. Read the data from an Excel file and load into a tibble
```{r Read data file(s)}

# change the Excel file name accordingly
df <- read_excel("data/dataset_verkopen.xlsx") # read Excel file and load it into data frame (change file name accordingly)

```
`End read and load data`

<BR><BR>

### 3. Data cleaning and preparation
```{r data cleaning and preparation}

# OPTIONAL: Remove empty observations (rows.)
# cleandf <- drop_na(exceldf, <NAME>) # Delete observations with the value NA after a specific variable
# cleandf <- na.omit(exceldf) # Remove all observations where a variable has the value NA

#knitr::opts_chunk$set(warning = FALSE, message = FALSE) # suppress all warning messages 

```
`End data cleaning and preparation`

\

***
**--> Example chunks <--**

***
\

### Example code chunk from dataframe
```{r Example histogram from dataframe}

  df %>% 
  select(Continent) %>% 
  ggplot(aes(Continent)) +
  geom_histogram(stat = "count")

```
`End example code chunk from dataframe`

\

***
**--> From here you make your own chunks <--**

***
\
