---
title: "R Notebook - Template - MySQL Data Access Using R (multiple tables)"
date: "March 2023"
author: "Daan Jansen"
output:
  html_document:
    df_print: paged
    html_notebook: default
    pdf_document: default
    theme:
      bootswatch: journal
---

\

***
We start with an extensive setup chunk in which we:

1. *Load* all the *packages* we use;
2. *Read* the *data* from a SQL Server database and load tables into separate tibbles
3. *Clean data* and preparation
\

***
\

### 1. Load libraries
```{r Loading libraries, include=FALSE}

# Packages have proprogrammed functions that solve specific problems and expands the R vocabulary
library("tidyverse") # includes packages like dplyr and ggplot2
library("lubridate") # working with dates and times
library("RMySQL") # database Interface and 'MySQL' Driver for R

```
`End load libraries`

<BR><BR><BR><BR>

### 2. Read the data from MySQL and load tables into separate tibbles
```{r loading from mssql, include=FALSE}

# Settings
db_host <- '127.0.0.1' # for local access or use studmysql01.fhict.local' when using FHICT Self Service Portal
db_port <- 3306
db_name <- 'database_name'
db_user <- 'your_name'
db_password <- 'your_password'
#db_user  <- rstudioapi::showPrompt('Login', 'Enter MySQL user name', '') # does not get saved to environment
#db_password  <- rstudioapi::askForPassword('Enter MySQL user password') # dont get saved to environment

# Read data from db
mydb <-  dbConnect(MySQL(), host = db_host, port = db_port, dbname = db_name, user = db_user, password = db_password)

# Store tables into dataframes
db_tables <- dbListTables(mydb)

# Iterate throught tables, load each table and assign it to a named dataframe
lapply(db_tables, function(x) {
  rs <- dbSendQuery(mydb, paste0("select * from ", x))
  assign(x, fetch(rs, n = -1), envir = sys.frame())
  rm(rs) # delete temporary variable
  })

#on.exit(dbDisconnect(mydb)) # close database connection

# Clean up
rm(db_user, db_password, db_name, db_host, db_port, db_tables)

```
`End read and load data'

<BR><BR>

### 3. Data cleaning and preparation
```{r data cleaning and preparation}

# OPTIONAL: Remove empty observations (rows.)
# cleandf <- drop_na(exceldf, <NAME>) # Delete observations with the value NA after a specific variable
# cleandf <- na.omit(exceldf) # Remove all observations where a variable has the value NA

#knitr::opts_chunk$set(warning = FALSE, message = FALSE) # suppress all warning messages 

#Information about tables present in database
#dbListTables(mydb)
#dbListFields(mydb, "<TABLE>") # change table name accordingly

```
`End data cleaning and preparation`

\

***
**--> Example chunks <--**

***
\

### Example SQL code chunk using dbGetQuery
```{r Example histogram from dbGetQuery}

# Returns the result of a SQL query as a data frame
query1 <- dbGetQuery(mydb,
                     "SELECT catcode, count(catcode) as aantal
                     FROM artikel
                     GROUP BY catcode")

query1 %>% 
  ggplot(aes(x = catcode, y = aantal)) + 
  geom_col()

```
`End example SQL code chunk using dbGetQuery`

<BR><BR>

### Example SQL code chunk capture in R dataframe and plot result
```{sql connection=mydb, output.var = "query1", label='Store result in outputvariable'}

SELECT catcode, count(catcode) as aantal
FROM artikel
GROUP BY catcode

```

```{r Plot result from previous SQL chunk}

query1 %>% 
  ggplot(aes(x = catcode, y = aantal)) + 
  geom_col()

```
`End example SQL code chunk capture in R dataframe and plot result`

<BR><BR>

### Example code chunk from dataframe
```{r Example histogram from dataframe}

  artikel %>% 
  select(catcode) %>% 
  ggplot(aes(catcode)) +
  geom_histogram(stat = "count")

```
`End example code chunk from dataframe`

\

***
**--> From here you make your own chunks <--**

***
\
