---
title: "R Notebook - Template - MSSQL Data Access Using R (multiple tables)"
date: "March 2023"
author: "Daan Jansen"
output:
  html_document:
    df_print: paged
    html_notebook: default
    pdf_document: default
    theme:
      bootswatch: journal
---

\

***
We start with an extensive setup chunk in which we:

1. *Load* all the *packages* we use;
2. *Read* the *data* from a SQL Server database and load tables into separate tibbles
3. *Clean data* and preparation
\

***
\

### 1. Load libraries
```{r Loading libraries, include=FALSE}

# Packages have proprogrammed functions that solve specific problems and expands the R vocabulary
library("tidyverse") # includes packages like dplyr and ggplot2
library("lubridate") # working with dates and times
library("odbc") # DBI-compliant interface to Open Database Connectivity (ODBC) drivers

```
`End load libraries`

<BR><BR>

### 2. Read the data from a SQL Server database and load tables into separate tibbles
```{r loading from mssql, include=FALSE}

#sort(unique(odbcListDrivers()[[1]])) # list of available odbc drivers

# Settings
driver.name <- "ODBC Driver 18 for SQL Server"
server.name <- "your_server"
server.port <- 1433
db.name     <- "database_name"
user.name   <- "your_name"
user.pwd    <- "your_password"
#user.name   <- rstudioapi::showPrompt('Login', 'SQL Server User ID', '') # does not get saved to environment
#user.pwd    <- rstudioapi::askForPassword('SQL Server Password') # dont get saved to environment

# Read data from db
mydb <- dbConnect(odbc::odbc(), driver = driver.name,
                  server = server.name, port = server.port, database = db.name,
                  uid = user.name, pwd = user.pwd)

# Store tables into dataframes
db.tables <- dbListTables(mydb, schema="dbo")

# Iterate throught tables, load each table and assign it to a named dataframe
lapply(db.tables, function(x) {
  rs <- dbSendQuery(mydb, paste0("select * from ", x))
  assign(x, dbFetch(rs, n = -1), envir = sys.frame())
  rm(rs) # delete temporary variable
  })

#on.exit(dbDisconnect(mydb)) # close database connection

# Clean up
rm(driver.name, server.name, server.port, db.name, user.name, user.pwd, db.tables)

```
`End read and load data`

<BR><BR>

### 3. Data cleaning and preparation
```{r data cleaning and preparation}

# OPTIONAL: Remove empty observations (rows.)
# cleandf <- drop_na(exceldf, <NAME>) # Delete observations with the value NA after a specific variable
# cleandf <- na.omit(exceldf) # Remove all observations where a variable has the value NA

#knitr::opts_chunk$set(warning = FALSE, message = FALSE) # suppress all warning messages 

#Information about tables present in database
#dbListTables(mydb, schema="dbo")
#dbListFields(mydb, scheme="dbo", "<TABLE>") # change table name accordingly

```
`End data cleaning and preparation`

\

***
**--> Example chunks <--**

***
\

### Example SQL code chunk using dbGetQuery
```{r Example histogram from dbGetQuery}

# Returns the result of a SQL query as a data frame
query1 <- dbGetQuery(mydb,
                     "SELECT catcode, count(catcode) as aantal
                     FROM artikel
                     GROUP BY catcode")

query1 %>% 
  ggplot(aes(x = catcode, y = aantal)) + 
  geom_col()

```
`End example SQL code chunk using dbGetQuery`

<BR><BR>

### Example SQL code chunk capture in R dataframe and plot result
```{sql connection=mydb, output.var = "query1", label='Store result in outputvariable'}

SELECT catcode, count(catcode) as aantal
FROM artikel
GROUP BY catcode

```

```{r Plot result from previous SQL chunk}

query1 %>% 
  ggplot(aes(x = catcode, y = aantal)) + 
  geom_col()

```
`End example SQL code chunk capture in R dataframe and plot result`

<BR><BR>

### Example code chunk from dataframe
```{r Example histogram from dataframe}

  artikel %>% 
  select(catcode) %>% 
  ggplot(aes(catcode)) +
  geom_histogram(stat = "count")

```
`End example code chunk from dataframe`

\

***
**--> From here you make your own chunks <--**

***
\
